Peer assignment #1
========================================================

This markdown was created as a exercise for the peer assignment 1 of the course Reproducible 
Research on Coursera.

17.1.2015
Author: Alena D.

### Loading and preprocessing the data

Datasets for this assignment contain personal activity information during 2 months in the 5 minutes 
intervals. The variables included in dataset are:
- steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
- date: The date on which the measurement was taken in YYYY-MM-DD format
- interval: Identifier for the 5-minute interval in which measurement was taken


#### Loading:
```{r}
data <- read.csv("d:/online_kurzy/Coursera/reproducible_research/peerass1/activity.csv")
```


#### Structure of uploaded data:
```{r}
str(data)
```

The proccessing steps are included in parts answering quenstions from assignment.


### What is mean total number of steps taken per day?

#### Processing data
```{r}
totStepsDays <- aggregate(steps ~ date,data,FUN=sum)
```

#### Histogram of total number of steps taken each day
```{r}
hist(totStepsDays$steps, main="Histogram of total number of steps per day", 
     xlab="# of steps per day")
```

#### Mean of total number of steps taken each day
```{r}
mean(totStepsDays$steps)
```
The mean number of steps per day is 10766.

#### Median of total number of steps taken each day
```{r}
median(totStepsDays$steps)
```
The median number of steps per day is 10765.



### What is the average daily activity pattern?

#### Processing data
```{r}
actPattern <- aggregate(steps ~ as.factor(data$interval),data,FUN=mean)
names(actPattern) <- c("int", "steps")
actPattern$int <- as.numeric(as.character(actPattern$int))
```

#### Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
library(ggplot2)
ggplot(data=actPattern, aes(x=int, y=steps)) + geom_line() + xlab("Minutes") + 
  ylab("Average number of steps accross all days") + theme_bw() +
  ggtitle("Average number of steps accross all days in every 5 minute interval")

```

### Imputing missing values

#### Number of missing values
```{r}
nrow(data[is.na(data)[,1],])
```
In primary data there are 2304 rows which contain NA value.

#### Strategy for filling in all of the missing values in the dataset and new dataset without missing values
```{r}
datawithoutNA <- data
datawithoutNA[is.na(datawithoutNA)[,1],1] <- actPattern[actPattern$int == datawithoutNA[is.na(datawithoutNA)[,1],3],2]
```
I selected a method for filling in NA values in using mean for that 5 minute interval.

#### Consequences of data manipulation
```{r}
totStepsDays <- aggregate(steps ~ date, datawithoutNA, FUN=sum)
hist(totStepsDays$steps, main="Histogram of total number of steps per day", 
     xlab="# of steps per day")
mean(totStepsDays$steps)
median(totStepsDays$steps)
```

The mean and median of total steps per day are equal after the data manipulation. Median was 10765 before.

### Are there differences in activity patterns between weekdays and weekends?

```{r}
data <- cbind(data, c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", 
  "Friday", "Saturday")[as.POSIXlt(data$date)$wday + 1])
data[,4] <- as.character(data[,4])
data[-grep("S", data[,4]),4] <- "weekday"
data[grep("S", data[,4]),4] <- "weekend"
data[,4] <- as.factor(data[,4])
str(data)
```

#### The 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
```{r}
actPattern <- aggregate(steps ~ as.factor(data$interval) + data[,4],data,FUN=mean)
names(actPattern) <- c("int", "dayw", "steps")
actPattern$int <- as.numeric(as.character(actPattern$int))

ggplot(data=actPattern, aes(x=int, y=steps)) + geom_line() + xlab("Minutes") + 
  ylab("Average number of steps accross all days") + theme_bw() +
  ggtitle("Average number of steps accross all days in every 5 minute interval") + facet_grid(dayw ~ .)
```





